---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

1. Load the data
```{r}
act<-read.csv(unzip("activity.zip", "activity.csv"))    #load data
```

2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
str(act)            #date column is factor. need to change to date. step column has many NA 
summary(act)        #all positive value.

unique(is.na(act$interval))
unique(is.na(act$date))         #only step column has NA values

act$date<-as.Date(as.character(act$date))     #change to date from factor
```


## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day
```{r}
sum1<-tapply(act$steps, act$date, sum)
sum2<-data.frame(date=names(sum1), sum=sum1)    
#according to the total sums, 10/1, 10/8, 11/1, 11/4, 11/9, 11/10, 11/14, 11/30 has NA sum.
#needs to check if all steps value are NA for all the entries for these dates or if R produced NA for even just one NA value
library(dplyr)
NAs<-filter(act, date %in% as.Date(as.character(sum2[is.na(sum2$sum),1])))    # find all rows with the date's total sum equals NA
unique(NAs$steps)     #check if there is any other value than NA. It seems like all values are NA.

#Total number of steps take per day is
sum2<-sum2[!is.na(sum2$sum),]
sum2


```


2. Make a histogram of the total number of steps taken each day

```{r}
hist(sum2$sum,main="Total step histogram", xlab="Total steps per day")
```

3. Calculate and report the mean and median of the total number of steps taken per day

```{r}
summary(sum2)     #mean=10766 median=10765
```


## What is the average daily activity pattern?

1. Make a time series plot (i.e.type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)


2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
