---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data


1. Load the data
```{r}
act<-read.csv(unzip("activity.zip", "activity.csv"))    #load data
```

2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
str(act)            #date column is factor. need to change to date. step column has many NA 
summary(act)        #all positive value.

unique(is.na(act$interval))
unique(is.na(act$date))         #only step column has NA values

act$date<-as.Date(as.character(act$date))     #change to date from factor
```


## What is mean total number of steps taken per day?


1. Calculate the total number of steps taken per day
```{r}
sum1<-tapply(act$steps, act$date, sum)
sum2<-data.frame(date=names(sum1), sum=sum1)    
#according to the total sums, 10/1, 10/8, 11/1, 11/4, 11/9, 11/10, 11/14, 11/30 has NA sum.
#needs to check if all steps value are NA for all the entries for these dates or if R produced NA for even just one NA value
library(dplyr)
NAs<-filter(act, date %in% as.Date(as.character(sum2[is.na(sum2$sum),1])))    # find all rows with the date's total sum equals NA
unique(NAs$steps)     #check if there is any other value than NA. It seems like all values are NA.

#Total number of steps take per day is
sum2<-sum2[!is.na(sum2$sum),]
sum2


```


2. Make a histogram of the total number of steps taken each day

```{r}
hist(sum2$sum,main="Total step histogram", xlab="Total steps per day")
```

3. Calculate and report the mean and median of the total number of steps taken per day

```{r}
summary(sum2)     #mean=10766 median=10765
```


## What is the average daily activity pattern?

1. Make a time series plot (i.e.type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
library(ggplot2)
int_mean<-act %>% group_by(interval) %>% summarize(mean=mean(steps, na.rm=TRUE))  #mean steps by interval
ggplot(data=int_mean, type="l",aes(x=interval, y=mean))+geom_line(size=1)+labs(x="min interval", y="mean across all days", title="Average steps by 5-min interval")

```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
int_mean[which.max(int_mean$mean),1] #835 min interval
```

## Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
nrow(act[is.na(act$steps),])  #2304 NA value
```


2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.



3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
NAs_date<-unique(NAs$date)
act_r<-act
NA_r<-which(is.na(act$steps))
for (i in 1:length(NA_r)) {
  act_r[NA_r[i],1]=int_mean[which(int_mean$interval==act[NA_r[i],]$interval),2] #replace all NA values with mean interval for all days
}
```




4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?


## Are there differences in activity patterns between weekdays and weekends?
